
#LOAD LIBRARIES
```{r}
library(tinytex)
library(knitr)
library(dplyr)
library(stringr)
library(caret)
library(randomForest) #RANDOM FOREST
library(pROC) #ROC CURVE
library(rpart) #DECISSION TREE
library(rpart.plot) #DECISSION TREE
library(e1071)
library(datasets)
library(class)
library(tidyr)
library(tidyverse)
library(broom)
library(AICcmodavg) #ANOVA feature selection
#clustering
library(cluster)
library(ggplot2)
library(factoextra)
library(dendextend)
library(stats)
library(graphics)
```










#LOAD THE DATASET
```{r}
train<-read.csv("H:/Windows Folders/Downloads/train.csv")
test<-read.csv("H:/Windows Folders/Downloads/test.csv") #There is no prediction column for validation hence we will not be using this dataset
sum(is.na(train))
```










#TREATMENT FOR MISSING VALUES
```{r}
head(train)

#Removing the unique ID column 
train<-train[-c(1)]

Tot_NA <- sum(is.na(train))/prod(dim(train)) #proportion of missing values

Col_NA <- sort(colMeans(is.na(train)), decreasing=TRUE) #Column NA Values
kable(Col_NA[which(Col_NA>0)], caption="Column NA Rates", col.names = c("NA Percentage")) #Column NA Values > 0%

#Remove NA Columns > Threshold
threshold_na = 0.4
Col_Remove = names(Col_NA[which(Col_NA>threshold_na)]) #These columns will be eliminated because the missing values are greater than 40% of the total entries in that field

#Columns which have missing values but below threhold
Col_Mean = names(Col_NA[which(Col_NA<=threshold_na & Col_NA>0)]) 
Col_Impute = names(Col_NA[which(Col_NA<=threshold_na & Col_NA>0)])

train <- train[,which(!names(train) %in% Col_Remove)] #Remove columns where NA exceeds the threshold value

#Removing the NA's in other columns - Examine whether they are categorical (MODE) or numerical (MEAN)
Col_Mean

head(train$Family_Hist_4) #numerical - mean
head(train$Employment_Info_6) #numerical - mean
head(train$Medical_History_1) #Categorical - mode
head(train$Employment_Info_4) #numerical - mean
head(train$Employment_Info_1) #numerical - mean

#since medical_history_1 is a categorical variable it will be replaced with the mode instead of the mean 
Col_Impute = Col_Impute[-c(3)] 
Col_Mean = Col_Mean[-c(3)]

#Mode Imputation Function
getmode <- function(v) {
   uniqv <- unique(v[!is.na(v)])
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

#Mode for missing NA's in categorical
train$Medical_History_1[which(is.na(train$Medical_History_1))] <- getmode(train$Medical_History_1)

#Mean for missing NA's in numerical
train$Family_Hist_4[which(is.na(train$Family_Hist_4))] <- mean(train$Family_Hist_4, na.rm = TRUE)
train$Employment_Info_6[which(is.na(train$Employment_Info_6))] <- mean(train$Employment_Info_6, na.rm = TRUE)
train$Employment_Info_4[which(is.na(train$Employment_Info_4))] <- mean(train$Employment_Info_4, na.rm = TRUE)
train$Employment_Info_1[which(is.na(train$Employment_Info_1))] <- mean(train$Employment_Info_1, na.rm = TRUE)

#Validate if all missing values are dealt with
sum(is.na(train))
dim(train)
```










#FEATURES ENGINEERING
```{r}
#Product_info_2 has been factorized into 2 columns thereby reducing the number of factors

train$Product_Info_2_char <- as.factor(substr(train$Product_Info_2, 1, 1))
head(train$Product_Info_2_char)
train$Product_Info_2_num <- as.factor(substr(train$Product_Info_2, 2, 2))
head(train$Product_Info_2_num)

str(train)
train<-train[-c(2)] #Once the 2 new columns are created, remove the original

train.base<-train #Backup train.base has been created

train$Med_keyword_count <- as.numeric(apply(train[,70:117], 1, sum)) #Adding the keywords
head(train$Med_keyword_count)
```











#FORMATTING THE DATA
```{r}
str(train)
names(train)
dim(train)

train[,1:2] <-lapply(train[,1:2],as.factor)
train[,4:6] <-lapply(train[,4:6],as.factor)
train[,c(12,13,15)] <-lapply(train[,c(12,13,15)],as.factor)
train[,17:31] <-lapply(train[,17:30],as.factor)
train[,33:117] <-lapply(train[,33:117],as.factor)
train[,c(33,34)] <-lapply(train[,c(33,34)],as.numeric)

str(train[,80:118]) #allows us to see till the very last column

train$Response<-as.factor(train$Response)

summary(train$Med_keyword_count)
#train$Med_keyword_count<-as.factor(train$Med_keyword_count) This can be converted to factor or left as integer. For now this has been left as integers
```











#CHECK THE DATA FOR OUTLIERS
```{r}
boxplot(train[1:10])$Response
boxplot(train[11:20])$Response
boxplot(train[21:30])$Response
boxplot(train[31:40])$Response
boxplot(train[41:50])$Response
boxplot(train[51:60])$Response
boxplot(train[61:70])$Response
boxplot(train[71:80])$Response
boxplot(train[81:90])$Response
boxplot(train[91:100])$Response
boxplot(train[101:110])$Response
boxplot(train[111:121])$Response

boxplot(train$Product_Info_2_char)$Response

boxplot(train$Product_Info_2_num)$Response #NOT AN OUTLIER - MANY RECORDS FOR CATEGORY 8
table(train$Product_Info_2_num)

boxplot(train$Medical_History_1)$Response #INVESTIGATE - there is some skewness observed here (Revisit after model fitting)
summary(train$Medical_History_1)
boxplot(train$Medical_History_2)$Response #INVESTIGATE - slight skewness can be seen (Revisit after model fitting) 
summary(train$Medical_History_2)

#NO MAJOR CONCERN WITH OUTLIER SO WE MOVE AHEAD
```











#DATA REDUCTION
```{r}
#Medical_keyword_1 to Medical_Keyword_48 has many 0's
#Typically indicating if any of those lives have the medical keywords in them
#instead of using that many columns with 0 values, we have added the values indicating the total number of medical keywords in each insured portfolio

names(train[70]) #Medical_Keyword_1
names(train[117]) #Medical_Keyword_48

train<-train[,-c(70:117)]
str(train)

#CHECK FOR MULTI-COLLINEARITY
cor(train[, unlist(lapply(train, is.numeric))]) #CHECK FOR MULTI-COLLINEARITY
#ONLY BMI AND Wt have a collinearity of 85. Hence, no column has been removed for now. 
```

#CHECK FOR DATA IMBALANCE 
```{r}
train$Response<-as.numeric(train$Response)
response_freq<-hist(train$Response,col = "steel blue")
response_freq
train$Response<-as.factor(train$Response)
```
As visible from the frequency diagram, class 8 (highest level of risk) seems to have the most amount of data and 3 (&4) has the least. There are two ways to correct for "DATA IMBALANCE" 
1. Combining 7+8 as "High Risk" & the others will be "Low Risk"
Since the magnitude of risk increases with each next category (0-8), intuitively it would be okay to combine 7&8 and classify that as "High-Risk" while the remaining categories would be "Low/Medium Risk". Now, converting the problem to a binary classification problem.  










#CONVERTING TO A BINARY PROBLEM
```{r}
train_binary<-train
train_binary$Response<-as.numeric(train_binary$Response)
summary(train_binary$Response)

train_binary$Response_binary <- ifelse(train_binary$Response > 6, 1, 0)
train_binary$Response_binary<-as.factor(train_binary$Response_binary)
summary(train_binary$Response_binary)

colnames(train_binary) #Response_binary is now the required response variable
names(train_binary)[70]
train_binary <- train_binary[,-c(70)] #delete the multi-class response variable

dim(train_binary)
```









#MERGING FACTORS WHICH HAVE A VERY FEW OBSERVATIONS INTO 1 FACTOR AS THIS MAY GIVE ERROR WHILE FITTING THE MODEL. 
Example - Few factors had only 1-2 observations. These observations can possibly be a part of the TEST DATA and not be present in the TRAIN DATA. This will throw an error while making out predictions. Hence this merging was carried out, and then we go on to split the data into TRAIN & TEST

```{r}
#MERGING FACTORS WITH VERY FEW OBSERVATIONS AS IT GIVES ERROR WHILE FITTING THE MODEL

Merge.factors <- function(x, p) { 
  #Combines factor levels in x that are less than a specified proportion, p.
  t <- table(x)                 
  less <- subset(t, prop.table(t) < p)
  more <- subset(t, prop.table(t) >= p)
  other <- rep("Other", sum(less))
  new.table <- c(more, table(other))
  new.x <- as.factor(rep(names(new.table), new.table))
  return(new.x)
}

table(train_binary$Product_Info_3)
train_binary$Product_Info_3<-Merge.factors(train_binary$Product_Info_3,0.005)


table(train_binary$Medical_History_17)
train_binary$Medical_History_17 <- ifelse(train_binary$Medical_History_17 == 1, 3,train_binary$Medical_History_17) 
train_binary$Medical_History_17<-as.factor(train_binary$Medical_History_17)

table(train_binary$Employment_Info_2)
train_binary$Employment_Info_2<-Merge.factors(train_binary$Employment_Info_2,0.005)

table(train_binary$Insurance_History_3)
train_binary$Insurance_History_3 <- ifelse(train_binary$Insurance_History_3 == 2, 3, train_binary$Insurance_History_3)
train_binary$Insurance_History_3<-as.factor(train_binary$Insurance_History_3)
```









#SPLIT THE DATA INTO TEST AND TRAIN WHICH WILL BE USED FOR MODEL BUILDING
```{r}

set.seed(1)
trainIndex_binary <- createDataPartition(train_binary$Response_binary, p = .8, 
                                  list = FALSE, 
                                  times = 1)

train.binary.data <- train_binary[ trainIndex_binary,]
test.binary.data  <- train_binary[-trainIndex_binary,]
```










#FIT A LOGISTIC REGRESSION MODEL TO ASSESS THE ACCURACY
```{r}
#FIT A LOGISTIC MODEL ON ALL THE DIMENSIONS
logit.reg <- glm(Response_binary ~ ., data = train.binary.data, family = "binomial") 
summary(logit.reg)

# use predict() with type = "response" to compute predicted probabilities. 
logit.reg.pred <- predict(logit.reg, test.binary.data, type = "response")
logit.reg.pred.train <- predict(logit.reg, train.binary.data, type = "response")

# first 5 actual and predicted records
data.frame(actual = test.binary.data$Response_binary[1:5], predicted = logit.reg.pred[1:5])

# classification
trainEstimatedResponse=ifelse(logit.reg.pred.train >0.5,1,0)
trainEstimatedResponse
testEstimatedResponse=ifelse(logit.reg.pred >0.5,1,0)
testEstimatedResponse


# Confusion matrix
#Accuracy, Estimation
table(train.binary.data$Response_binary, trainEstimatedResponse)
table(test.binary.data$Response_binary, testEstimatedResponse)

#Accuracy
mean(trainEstimatedResponse==train.binary.data$Response_binary)
mean(testEstimatedResponse==test.binary.data$Response_binary)

train.cm.log<-confusionMatrix(as.factor(ifelse(logit.reg.pred.train>0.5, 1, 0)), 
                train.binary.data$Response_binary)
test.cm.log<-confusionMatrix(as.factor(ifelse(logit.reg.pred>0.5, 1, 0)), 
                test.binary.data$Response_binary)
train.cm.log
test.cm.log

# create empty accuracy table
accT = c() 

# compute accuracy per cutoff
for (cut in seq(0,1,0.1)){
  cm <- confusionMatrix(as.factor(ifelse(logit.reg.pred>cut, 1, 0)), 
                        test.binary.data$Response_binary)
  accT = c(accT, cm$overall[1])
}
accT

# plot accuracy
plot(accT ~ seq(0,1,0.1), xlab = "Cutoff Value", ylab = "", type = "l", ylim = c(0, 1))
lines(1-accT ~ seq(0,1,0.1), type = "l", lty = 2)
legend("topright",  c("accuracy", "overall error"), lty = c(1, 2), merge = TRUE)

test_roc = roc(test.binary.data$Response_binary ~ logit.reg.pred, plot = TRUE, print.auc = TRUE)
```










#FIT A RANDOM FOREST MODEL AND ASSESS ACCURACY
```{r}
set.seed(1)

sqrt(dim(train.binary.data)[2]) #Value of mtry

rf <- randomForest(as.factor(Response_binary) ~ ., data = train.binary.data, ntree = 500, 
                   mtry = 9, nodesize = 5, importance = TRUE, parms = list(loss = lossmatrix))  

varImpPlot(rf, type = 1)

rf.pred <- predict(rf, test.binary.data)

test.cm.rf.1<-confusionMatrix(rf.pred, as.factor(test.binary.data$Response_binary))

#after removing the loss matrix
rf_2 <- randomForest(as.factor(Response_binary) ~ ., data = train.binary.data, ntree = 500, 
                   mtry = 9, nodesize = 5, importance = TRUE)  

varImpPlot(rf_2, type = 1)

rf.pred_2 <- predict(rf_2, test.binary.data)

test.cm.rf.2<-confusionMatrix(rf.pred_2, as.factor(test.binary.data$Response_binary))

test.cm.rf.1
test.cm.rf.2

rf.roc<-roc(train.binary.data$Response_binary,rf$votes[,2])
plot(rf.roc,print.auc = TRUE)
auc(rf.roc)
```









#FIT A DECISSION TREE MODEL AND ASSESS ACCURACY
```{r}
set.seed(1)

# classification tree
default.ct <- rpart(Response_binary ~ ., data = train.binary.data, method = "class")
length(default.ct$frame$var[default.ct$frame$var == "<leaf>"])
## Ans 4

# plot tree
prp(default.ct, type = 2, extra = 1, under = TRUE, split.font = 1, varlen = -10,  box.palette=c("red", "green"))

#### The full Tree
# argument cp sets the smallest value for the complexity parameter.
# argument  minsplit requires that the minimum number of observations in a node be 5 before attempting
deeper.ct <- rpart(Response_binary ~ ., data = train.binary.data, method = "class", cp = 0,minsplit = 30)

# count number of leaves
length(deeper.ct$frame$var[deeper.ct$frame$var == "<leaf>"])
## 1051

# plot tree
prp(deeper.ct, type = 1, extra = 1, under = TRUE, split.font = 1, varlen = -10,
    box.col=ifelse(deeper.ct$frame$var == "<leaf>", 'gray', 'white'),)  

#### Assessing the performance

# classify records in the test data.
# set argument type = "class" in predict() to generate predicted class membership.
default.ct.point.pred.train <- predict(default.ct,test.binary.data,type = "class")

# generate confusion matrix for test data
dt.cm.base.tree <- confusionMatrix(default.ct.point.pred.train, as.factor(test.binary.data$Response_binary))

### repeated the code for the deeper tree
deeper.ct.point.pred.test <- predict(deeper.ct,test.binary.data,type = "class")

# generate confusion matrix for test data
dt.cm.deeper.tree <- confusionMatrix(deeper.ct.point.pred.test, as.factor(test.binary.data$Response_binary))

### By default, rpart uses gini impurity to select splits when performing classification. 
#Now assessing the information gain instead by specifying it in the parms parameter.
IG.ct <- rpart(Response_binary ~ ., data = train.binary.data, method = "class", parms = list(split = 'information'))

# plot tree
prp(IG.ct, type = 2, extra = 1, under = TRUE, split.font = 1, varlen = -10,  box.palette=c("red", "green"))

IG.ct.pred.test <- predict(IG.ct,test.binary.data,type = "class")

# generate confusion matrix for test data
dt.cm.ig.tree <- confusionMatrix(IG.ct.pred.test, as.factor(test.binary.data$Response_binary))


### Now, if we were to penalize the more missclassified non-acceptor than acceptor then we can add a loss cost to the model as done below 
lossmatrix <- matrix(c(0,1,3,0), byrow = TRUE, nrow = 2)
lossmatrix

Loss.ct <- rpart(Response_binary ~ ., data = train.binary.data, method = "class", parms = list(loss = lossmatrix))
prp(Loss.ct, type = 2, extra = 1, under = TRUE, split.font = 1, varlen = -10,  box.palette=c("red", "green"))
Loss.ct.pred.test <- predict(Loss.ct,test.binary.data,type = "class")
# generate confusion matrix for test data
dt.cm.lm <- confusionMatrix(Loss.ct.pred.test, as.factor(test.binary.data$Response_binary))

#### cross-validation procedure
# argument xval refers to the number of folds to use in rpart's built-in
# cross-validation procedure
# argument cp sets the smallest value for the complexity parameter.
# argument  minsplit requires that the minimum number of observations in a node be 5 before attempting
#a split and that a split must decrease the overall lack of fit by a factor of 0.00001 (cost complexity factor) before being attempted.
#If the next best split in growing a tree does not reduce the tree’s overall complexity by 0.00001, rpart will terminate the growing process. 

cv.ct <- rpart(Response_binary ~ ., data = train.binary.data, method = "class", 
               cp = 0.00001, minsplit = 20)
prp(cv.ct, type = 2, extra = 1, under = TRUE, split.font = 1, varlen = -10,  box.palette=c("red", "green"))

length(cv.ct$frame$var[cv.ct$frame$var == "<leaf>"])
printcp(cv.ct)
plotcp(cv.ct)

rsq.rpart(cv.ct) # plot of approximate R-squared and relative error for different splits 
print(cv.ct)

# prune by lower cp
pruned.ct <- prune(cv.ct, 
                   cp = cv.ct$cptable[which.min(cv.ct$cptable[,"xerror"]),"CP"])
length(pruned.ct$frame$var[pruned.ct$frame$var == "<leaf>"])

prp(pruned.ct, type = 2, extra = 1, split.font = 1, varlen = -10, box.palette=c("red", "green"))  

pruned.ct.point.pred.train <- predict(pruned.ct,train.binary.data,type = "class")
prune.cm.train <- confusionMatrix(pruned.ct.point.pred.train, as.factor(train.binary.data$Response_binary))

pruned.ct.point.pred.test <- predict(pruned.ct,test.binary.data,type = "class")
prune.cm.test <-confusionMatrix(pruned.ct.point.pred.test, as.factor(test.binary.data$Response_binary))

#### prune by lower cp and considering std
set.seed(1)
cv.ct <- rpart(Response_binary ~ ., data = train.binary.data, method = "class", cp = 0.00001, minsplit = 1, xval = 5) 
length(cv.ct$frame$var[cv.ct$frame$var == "<leaf>"])

printcp(cv.ct) 
cv.ct.point.pred.test <- predict(cv.ct,test.binary.data,type = "class")
dt.cv.cm<-confusionMatrix(cv.ct.point.pred.test, as.factor(test.binary.data$Response_binary))

pruned.ct <- prune(cv.ct, cp = 0.000273)
prp(pruned.ct, type = 1, extra = 1, under = TRUE, split.font = 1, varlen = -10, 
    box.palette=c("red", "green")) 

pruned.ct.point.pred.test <- predict(pruned.ct,test.binary.data,type = "class")
dt.cv.prune.cm<- confusionMatrix(pruned.ct.point.pred.test, as.factor(test.binary.data$Response_binary))


dt.cm.base.tree
dt.cm.deeper.tree
dt.cm.ig.tree
dt.cm.lm
prune.cm.test
dt.cv.cm
dt.cv.prune.cm
```









#FIT A NAIVE BAYES MODEL AND ASSESS ACCURACY
```{r}
set.seed(1)  # Setting Seed

classifier <- naiveBayes(Response_binary ~ ., data = train.binary.data)

# Predicting on test data
y_pred <- predict(classifier, newdata = test.binary.data)
 
# Confusion Matrix
cm <- table(test.binary.data$Response_binary, y_pred)
nb.cm <- confusionMatrix(cm) 
nb.cm

nb_roc = roc(test.binary.data$Response_binary ~ as.numeric(y_pred), plot = TRUE, print.auc = TRUE)
```










#KNN MODEL AND ACCURACY
```{r}
set.seed(1)

#Converting the data to numeric so that KNN can be estimated correctly 
binary.KNN<- as.data.frame(sapply(train_binary, unclass)) #using the entire data for this and split later 
#Converts the factors/characters into numeric values so that we can find the nearest distance to assess the KNN model 

str(binary.KNN)

normalize <- function(x) {
  num <- x - min(x)
  denom <- max(x) - min(x)
  return (num/denom)
}

norm.binary.KNN<-as.data.frame(lapply(binary.KNN, normalize))
str(norm.binary.KNN)

#SPLIT THE NORMALIZED DATA
trainIndex.norm.KNN <- createDataPartition(norm.binary.KNN$Response, p = .8, 
                                  list = FALSE, 
                                  times = 1)

train.data.norm.KNN <- norm.binary.KNN[ trainIndex.norm.KNN,]
test.data.norm.KNN  <- norm.binary.KNN[-trainIndex.norm.KNN,]

colnames(train.data.norm.KNN)
names(test.data.norm.KNN)[73] #index of response variable

###########NORMALIZED DATA###########
# Build the model
travel_pred_k1 <- knn(train = train.data.norm.KNN[,-73], test = test.data.norm.KNN[,-73], cl = train.data.norm.KNN$Response_binary, k=1)
travel_pred_k5 <- knn(train = train.data.norm.KNN[,-73], test = test.data.norm.KNN[,-73], cl = train.data.norm.KNN$Response_binary, k=5)
travel_pred_k10 <- knn(train = train.data.norm.KNN[,-73], test = test.data.norm.KNN[,-73], cl = train.data.norm.KNN$Response_binary, k=10)
travel_pred_k20 <- knn(train = train.data.norm.KNN[,-73], test = test.data.norm.KNN[,-73], cl = train.data.norm.KNN$Response_binary, k=20)
travel_pred_k100 <- knn(train = train.data.norm.KNN[,-73], test = test.data.norm.KNN[,-73], cl = train.data.norm.KNN$Response_binary, k=100)
travel_pred_k200 <- knn(train = train.data.norm.KNN[,-73], test = test.data.norm.KNN[,-73], cl = train.data.norm.KNN$Response_binary, k=200)
travel_pred_k250 <- knn(train = train.data.norm.KNN[,-73], test = test.data.norm.KNN[,-73], cl = train.data.norm.KNN$Response_binary, k=250)

#Evaluation of Your Model
# Confusion matrix 
table(test.data.norm.KNN$Response_binary, travel_pred_k1)
table(test.data.norm.KNN$Response_binary, travel_pred_k5)
table(test.data.norm.KNN$Response_binary, travel_pred_k10)
table(test.data.norm.KNN$Response_binary, travel_pred_k20)
table(test.data.norm.KNN$Response_binary, travel_pred_k100)
table(test.data.norm.KNN$Response_binary, travel_pred_k200)
table(test.data.norm.KNN$Response_binary, travel_pred_k250)

knn.1.cm<-confusionMatrix(as.factor(test.data.norm.KNN$Response_binary), as.factor(travel_pred_k1))
knn.5.cm<-confusionMatrix(as.factor(test.data.norm.KNN$Response_binary), as.factor(travel_pred_k5))
knn.10.cm<-confusionMatrix(as.factor(test.data.norm.KNN$Response_binary), as.factor(travel_pred_k10))
knn.20.cm<-confusionMatrix(as.factor(test.data.norm.KNN$Response_binary), as.factor(travel_pred_k20))
knn.100.cm<-confusionMatrix(as.factor(test.data.norm.KNN$Response_binary), as.factor(travel_pred_k100))
knn.200.cm<-confusionMatrix(as.factor(test.data.norm.KNN$Response_binary), as.factor(travel_pred_k200))
knn.250.cm<-confusionMatrix(as.factor(test.data.norm.KNN$Response_binary), as.factor(travel_pred_k250))

# Classification accuracy 
sum(travel_pred_k1==test.data.norm.KNN$Response_binary)/length(test.data.norm.KNN$Response_binary)*100
sum(travel_pred_k5==test.data.norm.KNN$Response_binary)/length(test.data.norm.KNN$Response_binary)*100
sum(travel_pred_k10==test.data.norm.KNN$Response_binary)/length(test.data.norm.KNN$Response_binary)*100
sum(travel_pred_k20==test.data.norm.KNN$Response_binary)/length(test.data.norm.KNN$Response_binary)*100
sum(travel_pred_k100==test.data.norm.KNN$Response_binary)/length(test.data.norm.KNN$Response_binary)*100
sum(travel_pred_k200==test.data.norm.KNN$Response_binary)/length(test.data.norm.KNN$Response_binary)*100
sum(travel_pred_k250==test.data.norm.KNN$Response_binary)/length(test.data.norm.KNN$Response_binary)*100

#To study graphically which value of k gives us the best classification, we can plot Accuracy vs Choice of k.
# Empty variables
KNN_pred <- list()
accuracy <- numeric()

# From k=1 to k=100...
for(k in 1:100){
  # KnnTestPrediction for each k
  KNN_pred[[k]] <- knn(train.data.norm.KNN[,-73], test.data.norm.KNN[,-73], train.data.norm.KNN$Response_binary, k, prob=TRUE)
  
  # Accuracy for each k   
  accuracy[k] <- sum(KNN_pred[[k]]==test.data.norm.KNN$Response_binary)/length(test.data.norm.KNN$Response_binary)*100
  
}
# Accuracy vs Choice of k
?plot
plot(accuracy, type="b", col="dodgerblue", 
     xlab="k, number of neighbors", ylab="Classification accuracy", 
     main="Accuracy vs Neighbors")

# Add lines indicating k with best accuracy
abline(v=which(accuracy==max(accuracy)), col="orange", lwd=1.5)

# Add line for max accuracy seen
abline(h=max(accuracy), col="red", lty=2)
# lty=2 for dashed

# Add line for min accuracy seen 
abline(h=min(accuracy), col="grey", lty=2)

```










#USING CHI-SQUARE TEST AND ANOVA TEST FOR FEATURE SELECTION
After removing the variables which are not statistically significant to the response variable, we will assess the impact on the models.

```{r}
set.seed(1)

#CHI SQUARE FEATURE SELECTION (Categorical Vs Categorical)
str(test.binary.data)
colnames(train.binary.data)

select<-c(3,7,8,9,10,11,14,16,31,32,33,72)
print(chisq.test(train.binary.data[,3],train.binary.data$Response_binary))

CHIS <- lapply(train.binary.data[,-select], function(x) chisq.test(train.binary.data[,73], x))
chi_square_pvalue<-do.call(rbind, CHIS)[,c(1,3)]

summary(do.call(rbind, CHIS)[,c(3)]>0.05) #3 VARIABLES HAVE A HIGHER PValue which means they are not statistically significant
chi_square_pvalue[do.call(rbind, CHIS)[,c(3)]>0.05,] #These 3 categorical variables can be removed 
chi_square_selection<-rownames(chi_square_pvalue[do.call(rbind, CHIS)[,c(3)]>0.05,])

#Remove the above fields from the test and train data
test.post.chi.sq<-test.binary.data
test.post.chi.sq<-test.post.chi.sq[,!names(test.post.chi.sq) %in% chi_square_selection]
dim(test.post.chi.sq)

train.post.chi.sq<-train.binary.data
train.post.chi.sq<-train.post.chi.sq[,!names(train.post.chi.sq) %in% chi_square_selection]
dim(train.post.chi.sq)


#ANOVA FEATURE SELECTION (Numerical Vs Categorical)

names(train.binary.data)[select] #Numeric variables

#Only numeric variables have been considered (Numerical Vs categorical)
one.way.1 <- aov( Product_Info_4 ~ Response_binary, data = train.binary.data)
one.way.2 <- aov( Ins_Age ~ Response_binary , data = train.binary.data)
one.way.3 <- aov( Ht~ Response_binary, data = train.binary.data)
one.way.4 <- aov( Wt~ Response_binary, data = train.binary.data)
one.way.5 <- aov( BMI~ Response_binary, data = train.binary.data)
one.way.6 <- aov( Employment_Info_1~ Response_binary, data = train.binary.data)
one.way.7 <- aov( Employment_Info_4~ Response_binary, data = train.binary.data) 
one.way.8 <- aov( Employment_Info_6~ Response_binary, data = train.binary.data)
one.way.9 <- aov( Family_Hist_4~ Response_binary, data = train.binary.data)
one.way.10 <- aov( Medical_History_1~ Response_binary, data = train.binary.data)
one.way.11 <- aov( Med_keyword_count~ Response_binary, data = train.binary.data)
one.way.12 <- aov( Family_Hist_1~ Response_binary, data = train.binary.data)

options(scipen=9999)

summary(one.way.1)
summary(one.way.2)
summary(one.way.3)
summary(one.way.4)
summary(one.way.5)
summary(one.way.6)
summary(one.way.7)
summary(one.way.8)
summary(one.way.9)
summary(one.way.10)
summary(one.way.11)

#Since all numeric variables seem to be significant
test.post.chi.ano<-test.post.chi.sq
train.post.chi.ano<-train.post.chi.sq
```









#CHECK MODEL ACCURACY AFTER FEATURE ELIMINATION
#LOGISTIC REGRESSION
```{r}
set.seed(1)

logit.reg.chi <- glm(Response_binary ~ ., data = train.post.chi.ano, family = "binomial") 
summary(logit.reg.chi)

# use predict() with type = "response" to compute predicted probabilities. 
logit.reg.pred.chi <- predict(logit.reg.chi, test.post.chi.ano, type = "response")
logit.reg.pred.train.chi <- predict(logit.reg.chi, train.post.chi.ano, type = "response")

# first 5 actual and predicted records
data.frame(actual = test.post.chi.ano$Response_binary[1:5], predicted = logit.reg.pred.chi[1:5])

# classification
trainEstimatedResponse.chi=ifelse(logit.reg.pred.train.chi >0.5,1,0)
trainEstimatedResponse.chi
testEstimatedResponse.chi=ifelse(logit.reg.pred.chi >0.5,1,0)
testEstimatedResponse.chi


# Confusion matrix
#Accuracy, Estimation
table(train.post.chi.ano$Response_binary, trainEstimatedResponse.chi)
table(test.post.chi.ano$Response_binary, testEstimatedResponse.chi)

#Accuracy
mean(trainEstimatedResponse.chi==train.post.chi.ano$Response_binary)
mean(testEstimatedResponse.chi==test.post.chi.ano$Response_binary)

train.cm.log.chi<-confusionMatrix(as.factor(ifelse(logit.reg.pred.train.chi>0.5, 1, 0)), 
                train.post.chi.ano$Response_binary)
test.cm.log.chi<-confusionMatrix(as.factor(ifelse(logit.reg.pred.chi>0.5, 1, 0)), 
                test.post.chi.ano$Response_binary)
train.cm.log.chi
test.cm.log.chi
```

#RANDOM FOREST
```{r}
set.seed(1)

rf.chi <- randomForest(as.factor(Response_binary) ~ ., data = train.post.chi.ano, ntree = 500, 
                   mtry = 9, nodesize = 5, importance = TRUE, parms = list(loss = lossmatrix))  

varImpPlot(rf.chi, type = 1)

rf.pred.chi <- predict(rf.chi, test.post.chi.ano)

test.cm.rf.1.chi<-confusionMatrix(rf.pred.chi, as.factor(test.post.chi.ano$Response_binary))

#after removing the loss matrix
rf_2.chi <- randomForest(as.factor(Response_binary) ~ ., data = train.post.chi.ano, ntree = 500, 
                   mtry = 9, nodesize = 5, importance = TRUE)  

varImpPlot(rf_2.chi, type = 1)

rf.pred_2.chi <- predict(rf_2.chi, test.post.chi.ano)

test.cm.rf.2.chi<-confusionMatrix(rf.pred_2.chi, as.factor(test.post.chi.ano$Response_binary))

test.cm.rf.1.chi
test.cm.rf.2.chi

```










#CLUSTERING
```{r}
#test.post.chi.ano
#train.post.chi.ano

#Does the data have any impact in identifying the risk mapping?
#Let us look at the dependent/response field. Since this is the dependent variable, we can exclude this from our cluster analysis

#PROPORTION OF BINARY OUTPUT
p_0<-summary((train.post.chi.ano$Response_binary))[1]/(summary((train.post.chi.ano$Response_binary))[1]+summary((train.post.chi.ano$Response_binary))[2])

p_1<-summary((train.post.chi.ano$Response_binary))[2]/(summary((train.post.chi.ano$Response_binary))[1]+summary((train.post.chi.ano$Response_binary))[2])

#K-MEANS METHOD
set.seed(1)

#Data modifications
sum(is.na(train.post.chi.ano))
sapply(train.post.chi.ano, class)

train.post.chi.ano_cluster<-as.data.frame(sapply(train.post.chi.ano, unclass))

#Due to run computational limitations , we create a sample which can be used

clustering_sample<-sample_frac(train.post.chi.ano_cluster, 10000/nrow(train.post.chi.ano_cluster))

sapply(train.post.chi.ano_cluster, class)

#function to compute the within cluster - sum of squares for each value of k
wss<- function(k) {
  kmeans(train.post.chi.ano_cluster,k,nstart = 25)$tot.withinss
  }

k.values <- 1:10
wss_values <- map_dbl(k.values, wss)

plot(k.values, wss_values,
type="b", pch = 19, frame = FALSE,
xlab="Number of clusters K",
ylab="Total within-clusters sum of squares") #2 CLUSTERS SEEM TO BE THE BEST

# The other method with visual aid is using factoextra package
wss_sample<- function(k) {
  kmeans(clustering_sample,k,nstart = 25)$tot.withinss
  }
wss_values_sample <- map_dbl(k.values, wss_sample)

fviz_nbclust(clustering_sample, kmeans, method="wss") 
#Size was exceeding 13Gb hence a sample of 10000 records have been filtered to see the shape of elbow graph

#Average Silhouette Method:
fviz_nbclust(clustering_sample, kmeans, method = "silhouette")

#A merge
kmeans_perf2 = function(data,maxc,ns)
{
result = as.data.frame(matrix(ncol=3, nrow=maxc-1))
colnames(result) = c("clusters", "rsq","silhouette")
dst <- daisy(data)
for(i in 2:maxc) {
cst <- kmeans(data,i,iter.max=100,nstart=ns)
rsq <- 1-cst$tot.withinss/(cst$totss)
slht <- silhouette(cst$cluster,dst)
result[i-1,]=c(i,rsq,mean(slht[,3]))
}
ggplot(result, aes(clusters)) +
geom_line(aes(y = rsq, colour = "rsq")) +
geom_line(aes(y = silhouette, colour = "silhouette"))
}
kmeans_perf2(clustering_sample,10,ns=25) #GGPLOT

#The silhouette & merge methods confirm that the ideal number of clustering is 2.

#K = 2 SEEMS TO BE THE BEST CLUSTER COUNT
K2<-kmeans(train.post.chi.ano_cluster,2,nstart = 25)
K2
K2$cluster #CLUSTER 1 & 2 - WILL BE ADDED IN OUR DATA TO SEE IF THIS IMPROVES PREDICTION
names(K2)
K2$size

#clustering plot
fviz_cluster(K2, geom = "point", data = train.post.chi.ano_cluster) + ggtitle("k = 2")

#K-means clustering has embedded PCA which highlights two components accounting for only 12% of the variation
#in the overall demographic dataset. The k-means method does not provide a more granular segmentation than what the business already has

#K MEANS CLUSTERING FOR 2,3,4,5
K2
K3<-kmeans(train.post.chi.ano_cluster,3,nstart = 25)
K4<-kmeans(train.post.chi.ano_cluster,4,nstart = 25)
K5<-kmeans(train.post.chi.ano_cluster,5,nstart = 25)

cldata <- rbind(data.frame(clustering="kmeans 4",cluster=K5$cluster,pref=train.post.chi.ano_cluster$Response_binary),
data.frame(clustering="kmeans 3",cluster=K4$cluster,pref=train.post.chi.ano_cluster$Response_binary),
data.frame(clustering="kmeans 2",cluster=K3$cluster,pref=train.post.chi.ano_cluster$Response_binary),
data.frame(clustering="kmeans 2",cluster=K2$cluster,pref=train.post.chi.ano_cluster$Response_binary))
gp <- group_by(cldata,clustering,cluster)
print.data.frame(summarize(gp,size=n(),low=1-sum(pref==1)/length(pref),High=sum(pref==1)/length(pref)))

#WHILE THERE IS NO CLEAR STARC DIFFERENCE IN PROBABILITY OF HIGH & LOW RISK
#WE SEE RESULTS OF THE CLUSTER BEING IDENTIFIED WITH A PROBABILITY OF 54%-46% IN K MEANS 2,3,4. Since, the elbow plots have clearly indicated 2 as a good choice for the clusters, we will focus on k means 2.


#HIERARCHIEL CLUSTERING
#1.Elbow Method:
fviz_nbclust(clustering_sample, FUN = hcut, method = "wss")
#2.Average Silhouette Method:
fviz_nbclust(clustering_sample, FUN = hcut, method = "silhouette")
```

















#USE THE CLUSTERS GIVEN BY K-MEANS CLUSTERING TO ASSESS THE IMPACT ON PREDICTION
#LOGISTIC REGRESSION
```{r}
set.seed(1)
train_cluster <- train.post.chi.ano
train_cluster$cluster<-K2$cluster

#Clustering output for the test data which will act as a predictor
test.post.chi.ano_cluster<-as.data.frame(sapply(test.post.chi.ano, unclass))
K2_test<-kmeans(test.post.chi.ano_cluster,2,nstart = 25)

test_cluster <- test.post.chi.ano
test_cluster$cluster<-K2_test$cluster

train_cluster$cluster<-as.factor(train_cluster$cluster)
test_cluster$cluster<-as.factor(test_cluster$cluster)
dim(train_cluster)

#FIT A LOGISTIC MODEL ON ALL THE DIMENSIONS
logit.reg.cluster <- glm(Response_binary ~ ., data = train_cluster, family = "binomial") 
summary(logit.reg.cluster)

# use predict() with type = "response" to compute predicted probabilities. 
logit.reg.pred.cluster <- predict(logit.reg.cluster, test_cluster, type = "response")
logit.reg.pred.train.cluster <- predict(logit.reg.cluster, train_cluster, type = "response")

# first 5 actual and predicted records
data.frame(actual = test_cluster$Response_binary[1:5], predicted = logit.reg.pred.cluster[1:5])


# classification
trainEstimatedResponse.cluster=ifelse(logit.reg.pred.train.cluster >0.5,1,0)
testEstimatedResponse.cluster=ifelse(logit.reg.pred.cluster >0.5,1,0)

# Confusion matrix of KnnTestPrediction_k1
#Accuracy, Estimation
table(train_cluster$Response_binary, trainEstimatedResponse.cluster)
table(test_cluster$Response_binary, testEstimatedResponse.cluster)

#Accuracy
mean(trainEstimatedResponse.cluster==train_cluster$Response_binary)
mean(testEstimatedResponse.cluster==test_cluster$Response_binary)

#Confusion Matrix
train.cluster.cm<-confusionMatrix(as.factor(ifelse(logit.reg.pred.train.cluster>0.5, 1, 0)), 
                train_cluster$Response_binary)
test.cluster.cm<-confusionMatrix(as.factor(ifelse(logit.reg.pred.cluster>0.5, 1, 0)), 
                test_cluster$Response_binary)


test_roc = roc(test_cluster$Response_binary ~ logit.reg.pred.cluster, plot = TRUE, print.auc = TRUE)
```





#RANDOM FOREST (AFTER CHISQUARE & CLUSTERING)

```{r}
#train_cluster
#test_cluster
set.seed(1)

rf_cluster <- randomForest(as.factor(Response_binary) ~ ., data = train_cluster, ntree = 500, 
                   mtry = 4, nodesize = 5, importance = TRUE, parms = list(loss = lossmatrix))  

varImpPlot(rf_cluster, type = 1)

rf.pred.cluster <- predict(rf_cluster, test_cluster)

rf.cluster.cm<-confusionMatrix(rf.pred.cluster, as.factor(test_cluster$Response_binary))

#after removing the loss matrix
rf_cluster_2 <- randomForest(as.factor(Response_binary) ~ ., data = train_cluster, ntree = 500, 
                   mtry = 4, nodesize = 5, importance = TRUE)  

varImpPlot(rf_cluster_2, type = 1)

rf.pred_cluster_2 <- predict(rf_cluster_2, test_cluster)

rf.cluster.cm2<-confusionMatrix(rf.pred_cluster_2, as.factor(test_cluster$Response_binary))
```
